### 2024-03-12 rawfiles/letterhome.qmd
reorganising to tidy the repo look

---
title: "Draft open letter to editor"
format: html
editor: visual
---

To the editor and editorial board,

I am formally writing to point out the significant errors in the paper from Gibson (2024)[^1], published in your journal. In particular, the paper asserts two claims:

[^1]: Reference to a footnote.

-   That claims of negative New Zealand mortality are underpinned by the "Our World In Data" projection graph using the method of KRansky et all[^2].
-   That Gibson's method, producing postive excess mortality, is more effective than Kransky & in determining mortality as it adjusts for population change due to immigration changes.

[^2]: Kransky's article reference

Neither of these claims are accurate.

# The underpinng of New Zealand's mortality measurement

For centuries, it has been established practice among actuaries, demographers, and health researchers to measure mortality using death rates by age were such data is available[^3]. In New Zealand, tutorials on making mortality comparisons by using death rates by age and standardising to a reference population have been published since the 1890s\[\^NZ statistical handbook 1890\].

[^3]: 1767 whatever

Age Standardised Mortality direct adjusts for population growth and aging, as it is based on the number of deaths of the current resident population divided by the living current resident population, stratified by age group and time period. This directly measures the population change effects Gibson is trying to model, as any population change that does not change risk of dying leaves the rate of living people who die unchanged.

While calculation of New Zealand's cummulative excess mortality for 2020-2022 can vary by a couple of percentage points depending on the exact method, using a death rates by age (standardised) method the Institute of Actuaries find New Zealand's excess mortality from week 10 2020 to the end of 2022 to be -4% (that there are 4% fewer deaths than would be expected based on precovid trends)\[\^actuaries\].

The Our World in Data graph, and the K&K method that underlies it is an approximation, and while if it is a better or worse approximation to Gibson's method will be discussed later, while recognising the Our World In Data graph is an approximation there are still several practically useful implicit features about it.

-   Being based on trends of deaths, it enables a common basis comparison with countries that lack age stratified current population information
-   Weekly
-   Able to be checked by any member of the public for the latest current worldwide information. This last is a function of Our World in Data presentation rather than the method, but it is still a practical benefit.

While the existance of centuries of mortality measurement (not mentioned in Gibson's article) does not directly address if New Zealand medical professionals are aware that best practice mortality calculations recognise that people of different ages die at different rates when using the publicly available Our World in Data graphs in media pieces, it does establish that there is an established methodology, not mentiond in Gibson's article, underpinnning claims of New Zealand negative excess, and that method has been used for centuries, world-wide, across multiple disciplines. And we can check if New Zealand medical professionals are aware of this by asking them- " the key point being any comparisons both within country (over time) or between countries should be age standardised to provide the most accurate picture possible (and of course be based on high quality and complete underlying deaths data which we can say NZ's are!)" (Bloomfield, 2024) \[\^Pers Comm.\]

## Age Standardised Mortality Review

As Age Standardised Mortality is a standard across multiple disciplines, but is not raised in the Gibson paper, a brief review is in order.

Death rates for an age stratified group, for a time period in a given location, measure that rate at which that age group dies. These can be directly compared to death rates in other time periods and locations to compare if the group experience is more or less deadly. As it is a rate, it automatically adjusts for change in the population size of the age group.

While this enables direct comparisons of people of the same age, it does not enable total population comparisons (as total populations be may have different age structures). Standardisation with a reference population addresses this by assessing the total population deaths on the basis of death rates by age if the places and times being compared had the same age structure.

*Figure 1*

![ASMR steps](figure1.png){fig-alt="ASMR process" fig-align="center"}

While some variation in calculation results can stem from choice of standard population, due to the variation of young and old within a population, pragmatically any population that can be argued to be analytically useful should produce similar results. Using single year ages groups by sex to 94 then 95+, and a linear baseline from 2013-2019, a standard reference population of 2023 gives a cumulative excess of -5% for 2020 to 2022. Using a standard reference population of 2021 when the borders were restricted gives a cumulative excess of -6%. Using a pre-covid standard population of 2019 gives a cumulative excess of -5%. Using the standard population of 1961, used by StatsNZ to maintain long-term continuity, cumulative excess is -5%.

### Baselines and excess mortality

Excess mortality is the difference to what is expected for mortality, so excess calculations using any method are sensitive what is considered normal- the baseline. For Age Standardised Mortality the normal process is to use a linear trend of preceeding years, and in the case of a major multiyear event like a pandemic which does not yet have a post-acute pandemic trend established, the linear trend of pre-pandemic years.

As New Zealand, no different to most other advanced economies, has seen the rate of decrease in mortality slow over time, the length of the linear trend baseline period should be short enough not to be introducing error as a result of applying a straight line of best fit to curved data.

*Figure 2 goes here*

Based on the influenza season patterns in recent pre-pandemic years, the Australian Bureau of Statistics (ABS) considers 2013-2019 to give the best baseline\[\^ABS link\]. New Zealand has seen a very similar history in this period to Australia, with the added limitation that incorporating 2011, the year of the Christhchurch Earthquake and the residents of the second largest city living in broken housing through the following winter, is a poor addition to a baseline as it is not a re-occuring annual event.

While many international comparisons use a 5 year baseline range to increase the number of countries for which weekly level data is available, in practice the results for New Zealand do not differ substantively between a 5 (2015-2019) and 7 (2013-2019) year baseline period. A 5 year linear baseline with a 2022 standard population gives a cumulative excess for the period of -5%, while a 7 year baseline gives a cumulative baseline of -5%.

## Comparison with Gibson

With the existance of a common, internationally recognised best practice measure of mortality, age standised mortality, giving a cumulative excess mortality for 2020-2022 in the range of -4% to -6% depending on analytic assumptions.

The reason for Gibson's model preforming so poorly compared to